---
title: DANL 210 Project
subtitle: "Data Collection: Unifying Environmental, Social, and Governance (ESG) Metrics with Financial Analysis"
author: 
  - Dylan Thody
  - Owen Ellick
  - Jake Starkey

toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true

fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false

from: markdown+emoji
---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
install.packages("reticulate", repos = "https://cloud.r-project.org")

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```



# Introduction

The integration of Environmental, Social, and Governance (ESG) metrics with financial analysis has become increasingly significant in recent years. ESG metrics provide a comprehensive view of a company's ethical impact and sustainability practices, which are crucial for investors, regulators, and other stakeholders. Traditional financial analysis, while essential, often fails to capture these non-financial factors that can materially impact a company's performance and long-term value. Therefore, unifying ESG metrics with financial analysis is not only relevant but also necessary to foster sustainable economic growth, enhance risk management, and promote responsible investment practices. Despite the recognized importance of ESG factors, there is still a lack of standardization and integration in how these metrics are assessed and incorporated into financial analysis. This project aims to address the specific problem of fragmented and inconsistent ESG reporting, which creates challenges for investors and other stakeholders in making informed decisions. The primary issue is the absence of a unified framework that combines ESG metrics with traditional financial indicators, leading to gaps in the evaluation of a company's overall performance and sustainability. By developing a comprehensive and standardized approach to ESG and financial analysis integration, this research seeks to enhance the reliability and comparability of ESG data, ultimately supporting more effective and responsible decision-making in the financial markets.


# Data Collection

## yfinance

yfinance is a Python library that provides an easy to use interface in order to access financial data from Yahoo Finance. 


To use it, you have to first install yfinance using:

'pip install yfinance'


Then you create a ticker object for the company you are interested in using. For example I will create a ticker object for Apple below:

'ticker = yf.Ticker("AAPL")'


To retrieve financials for annual data, you use:

'ticker.financials'


To get access to a company balance sheet, you use:

'ticker.balance_sheet'


To get access to a Cash Flow Statement, you use:

'ticker.cashflow'


yfinance is highly effecient for pulling structured financial data quickly, making it ideal for financial anlaysis and modeling.

## selenium

selenium is a powerful tool for web scraping and browser automation. It can be used to extract ESG (Environmental, Social, and Governance) data from websites like Yahoo Finance that do not provide an API for such data.


To install selenium you use:

'pip install selenium'


To navigate to Web Pages you use the WebDriver to the target URL which in this case would be to the Yahoo Finance page.


Use WebDriver methods to locate and extract ESG data elements from the webpage using techniques such as XPath or CSS selectors.


You can also handle page loading and dynamic content using:

'time.sleep()'


Combining yfinance for financial and accounting data with selenium for ESG data allows you to comprehensively analyze a company’s performance and sustainability. This dual approach leverages the strengths of both libraries, providing a robust solution for data retrieval in financial analysis and ESG reporting.


# Descriptive Statistics

Below I am going to provide both grouped and un-grouped descriptive statistics and distribution plots for the ESG data and the finance/accounting data. In addition to that, I will provide correlation heat maps using cor() and seaborn.heatmap().

First you need to load the necessary libraries.

```{r}
library(reticulate)

# Set the Python path
use_python("/opt/anaconda3/bin/python3")

# Install required Python packages using reticulate
py_install("pandas")
py_install("seaborn")
py_install("matplotlib")

```

Then, you need to load and preprocess data using Python.

```{r}
library(reticulate)
use_python("/opt/anaconda3/bin/python3")

# Run Python code to load and preprocess data
py_run_string("
import yfinance as yf
import pandas as pd

# Load financial data
financial_data = pd.read_csv('yfinance-history.csv')

# Load ESG data
esg_data = pd.read_csv('esg_proj_data.csv')
esg_data.dropna(inplace=True)
esg_data.to_csv('cleaned_esg_data.csv', index=False)
")

# Read the data back into R
financial_data <- read.csv("financial_data.csv")
esg_data <- read.csv("cleaned_esg_data.csv")

```

Next up, I am going to make the descriptive statistics and distribution plots.

```{r}
# Ungrouped descriptive statistics for financial data
financial_desc <- summary(financial_data)

# Ungrouped descriptive statistics for ESG data
esg_desc <- summary(esg_data)

# Print descriptive statistics
print(financial_desc)
print(esg_desc)
```

```{r}
# Grouped descriptive statistics by company symbol for financial data
library(dplyr)
grouped_financial_desc <- financial_data %>%
  group_by(Symbol) %>%
  summarize(across(everything(), list(mean = mean, sd = sd)))

# Grouped descriptive statistics by company for ESG data
grouped_esg_desc <- esg_data %>%
  group_by(Comapny.Name) %>%
  summarize(across(everything(), list(mean = mean, sd = sd)))

# Print grouped descriptive statistics
print(grouped_financial_desc)
print(grouped_esg_desc)
```

Then, you calculate the descriptive statistics

```{r}
use_python("/opt/anaconda3/bin/python3")

# Run Python code to retrieve financial data using yfinance and preprocess ESG data
py_run_string("
import yfinance as yf
import pandas as pd

# Define the list of company symbols
companies = ['AAPL', 'MSFT', 'GOOGL', 'AMZN', 'FB']  # Example companies

# Create an empty DataFrame to store financial data
financial_data = pd.DataFrame()

for symbol in companies:
    ticker = yf.Ticker(symbol)
    financial_info = ticker.history(period='1y')[['Close']]
    financial_info['Symbol'] = symbol
    financial_data = pd.concat([financial_data, financial_info])

# Reset index for better handling
financial_data.reset_index(inplace=True)
financial_data.set_index(['Date', 'Symbol'], inplace=True)
financial_data.to_csv('financial_data.csv')

# Load ESG data
esg_data = pd.read_csv('esg_proj_data.csv')
esg_data.dropna(inplace=True)
esg_data.to_csv('cleaned_esg_data.csv', index=False)
")

# Read the data back into R
financial_data <- read.csv("financial_data.csv")
esg_data <- read.csv("cleaned_esg_data.csv")

# Ungrouped descriptive statistics
financial_desc <- summary(financial_data)
esg_desc <- summary(esg_data)

# Print ungrouped descriptive statistics
print("Ungrouped Descriptive Statistics for Financial Data:")
print(financial_desc)
print("Ungrouped Descriptive Statistics for ESG Data:")
print(esg_desc)

# Grouped descriptive statistics
grouped_financial_desc <- financial_data %>%
  group_by(Symbol) %>%
  summarize(across(where(is.numeric), list(mean = mean, sd = sd)), .groups = 'drop')

grouped_esg_desc <- esg_data %>%
  group_by(Comapny.Name) %>%
  summarize(across(where(is.numeric), list(mean = mean, sd = sd)), .groups = 'drop')

# Print grouped descriptive statistics
print("Grouped Descriptive Statistics for Financial Data:")
print(grouped_financial_desc)
print("Grouped Descriptive Statistics for ESG Data:")
print(grouped_esg_desc)

```

Then you create the distribution plots

```{r}
py_run_string("
import matplotlib.pyplot as plt
import seaborn as sns

# Load data
financial_data = pd.read_csv('financial_data.csv')
esg_data = pd.read_csv('cleaned_esg_data.csv')

# Distribution plot for financial data
plt.figure(figsize=(10, 6))
sns.histplot(financial_data['Close'], kde=True)
plt.title('Distribution of Closing Prices')
plt.xlabel('Close')
plt.ylabel('Frequency')
plt.show()

# Distribution plot for ESG data
plt.figure(figsize=(10, 6))
esg_columns = esg_data.columns[1:]  # Assuming first column is company name
for col in esg_columns:
    sns.histplot(esg_data[col], kde=True, label=col, element='step')
plt.title('Distribution of ESG Metrics')
plt.xlabel('Score')
plt.ylabel('Frequency')
plt.legend()
plt.show()
")
```

Then the correlation heat maps

```{r}
# Install and load necessary packages
library(reticulate)

# Set the Python path
use_python("/opt/anaconda3/bin/python3")
# Run Python code to create correlation heatmaps
py_run_string("
import seaborn as sns
import matplotlib.pyplot as plt
import pandas as pd

# Load data
financial_data = pd.read_csv('financial_data.csv')
esg_data = pd.read_csv('cleaned_esg_data.csv')

# Remove non-numeric columns for correlation calculation
financial_data_numeric = financial_data.select_dtypes(include=['float64', 'int64'])

# Correlation heatmap for financial data
plt.figure(figsize=(10, 8))
financial_corr = financial_data_numeric.corr()
sns.heatmap(financial_corr, annot=True, cmap='coolwarm', vmin=-1, vmax=1)
plt.title('Correlation Heatmap for Financial Data')
plt.show()

# Correlation heatmap for ESG data
plt.figure(figsize=(10, 8))
esg_corr = esg_data.iloc[:, 1:].corr()  # Assuming first column is company name
sns.heatmap(esg_corr, annot=True, cmap='coolwarm', vmin=-1, vmax=1)
plt.title('Correlation Heatmap for ESG Data')
plt.show()
")

```


# Exploratory Data Analysis

## Question 1

What are the mean and standard deviation of the closing prices for each company over the last year?

```{r}
library(reticulate)

use_python("/opt/anaconda3/bin/python3")

py_run_string("
import pandas as pd

financial_data = pd.read_csv('financial_data.csv')

mean_std = financial_data.groupby('Symbol')['Close'].agg(['mean', 'std'])
print(mean_std)
")

```

## Question 2

What is the distribution of closing prices for each company?

```{r}
library(reticulate)

use_python("/opt/anaconda3/bin/python3")

py_run_string("
import pandas as pd
import seaborn as sns
import matplotlib.pyplot as plt

financial_data = pd.read_csv('financial_data.csv')

plt.figure(figsize=(14, 7))
sns.histplot(data=financial_data, x='Close', hue='Symbol', multiple='stack', kde=True)
plt.title('Distribution of Closing Prices for Each Company')
plt.xlabel('Close')
plt.ylabel('Frequency')
plt.show()
")
```

## Question 3

Are there any significant correlations between the closing prices of different companies?


```{r}
library(reticulate)

use_python("/opt/anaconda3/bin/python3")

py_run_string("
import pandas as pd
import seaborn as sns
import matplotlib.pyplot as plt

financial_data = pd.read_csv('financial_data.csv')

financial_data_numeric = financial_data.pivot_table(index='Date', columns='Symbol', values='Close')

financial_corr = financial_data_numeric.corr()
print(financial_corr)

plt.figure(figsize=(10, 8))
sns.heatmap(financial_corr, annot=True, cmap='coolwarm', vmin=-1, vmax=1)
plt.title('Correlation Heatmap for Closing Prices')
plt.show()
")
```

##Question 4

How do the distributions of the Environmental, Social, and Governance scores compare?


```{r}
library(reticulate)

use_python("/opt/anaconda3/bin/python3")

py_run_string("
import pandas as pd
import seaborn as sns
import matplotlib.pyplot as plt

esg_data = pd.read_csv('cleaned_esg_data.csv')

plt.figure(figsize=(14, 7))
for col in esg_data.columns[1:]:
    sns.kdeplot(esg_data[col], label=col, fill=True)
plt.title('Distribution of ESG Scores')
plt.xlabel('Score')
plt.ylabel('Density')
plt.legend()
plt.show()
")
```


# Signifigance of the Project

Publishing a data analysis project on a publicly accessible platform like GitHub, focused on "Unifying Environmental, Social, and Governance (ESG) Metrics with Financial Analysis," promotes transparency, enhances credibility, and contributes to the body of knowledge. It supports informed decision-making by providing a holistic view of company performance, raising awareness about ESG's impact on financial performance, and encouraging the adoption of best practices. Hosting on GitHub leverages its features for version control and collaboration, facilitating peer review and ongoing improvements. By making your work available, you establish yourself as a knowledgeable entity, attract attention from potential employers and collaborators, and provide an educational resource for students and professionals.

# References

Collaboraters:
- Owen Ellick
- Jake Starkey

ChatGPT helped guide my codes

yahoo finance: https://finance.yahoo.com/



